Thermodynamic fragments for magic states in quantum computation

### 2021-03-22 magic_paper.tex
Tidied up unital and stabilizer fragment bound sections

\nick{We can skip mana and just discuss sum-negativity}
One of the most fundamental and commonly used magic monotones is the \emph{mana} of a state~\cite{cit:veitch2}, defined as
\begin{equation}
    \mana{\rho} \coloneqq \ln{(2\hspace{1pt}\sn{\rho}+1)},
\end{equation}
where the \emph{sum-negativity} ($\rm{sn}$)~\cite{cit:veitch2} is the sum of the negative components in $\W{\rho}$,
\begin{equation}
    \sn{\rho} \coloneqq \sum\limits_{\bmx: \W[\bmx]{\rho} < 0} \abs{\W[\bmx]{\rho}}.
\end{equation}
Mana is an additive\footnote{It satisfies the condition $\mana{\rho_1 \otimes \rho_2} = \mana{\rho_1} + \mana{\rho_2}$ which is practical in distillation settings.} magic monotone, so it provides an analytical, necessary condition for many-copy magic state interconvertibility.

@@

Note that the Wigner components of an $n$-copy state $\rho^{\otimes n}$ can be calculated directly from $\W{\rho}$ by convolution of the distribution with itself,
\begin{equation}
	\W{\rho^{\otimes n}} = \W{\rho}^{\otimes n},
\end{equation}
where $\otimes$ can be interpreted as the usual Kronecker product in the last expression.
We may use the vector notation $\bmw(\rho) = (w(\rho)_i)_{i=1,\dots,d^2}$ for the Wigner distribution, in which case we can express the Kronecker product as
\begin{equation}
	\W{\rho} \otimes \W{\rho} = (w(\rho)_i w(\rho)_j)_{i,j=1,\dots,d^2}.
\end{equation}
The correspondence between the phase space and pure vector representations of the Wigner distribuition is discussed more in~\cref{app:cmpairs}.

@@

Normalisation of the Wigner distribution \nick{REF} ensures that for all quantum states $\rho$, $\lc{\rho}{\sigma}(x) \geq 0$ and $\lc{\rho}{\sigma}(1) = 1$.
We stress that $0 \leq \lc{\rho}{\sigma}(x) \leq 1$ for all $x \in [0,1]$ if and only if $\rho$ is a positive Wigner state.
As a consequence, checking whether a magic state conversion of the form
\begin{equation}\label{eq:sigma_conv}
	\rho \xrightarrow{\E \in \O_\sigma} \tau
\end{equation} 
is not possible, reduces to the set of constraints
\begin{equation}\label{eq:majbound}
    \lc{\rho}{\sigma}(x) \geq \lc{\tau}{\sigma}(x),\ x\in [0,1],
\end{equation}
due to~\cref{thm:sigmamajor,thm:dmajor}.

We can refine the number of independent constraints stemming from this inequality. 
In fact, there are only as many independent constraints as there are elbows in the Lorenz curve of the target state as shown in~\cref{app:lc_constraints}.
However, in principle any one inequality defined at some location $x$ provides a valid constraint leading to some upper distillation bound, and the more locations one considers, the stricter the bound is.

We can now show that the magic monotone of mana, and equivalently the sum-negativity of a quantum state, can be viewed as a feature of majorisation, and independent of the particular $\sigma$--fragment one works in.
In fact, our current setting makes it apparent that mana provides a weaker condition than majorisation for all magic state interconversions.

@@

We can therefore view mana as just one feature of the Lorenz curve, namely its maximum value. 
Conversely, it is now clear that the maximum of the Lorenz curve acts as a valid magic monotone.
This implies the following result.
\begin{theorem}\label{thm:bounds}
    Given a magic state conversion $\rho \longrightarrow \tau$, the majorisation condition is stronger than the mana condition in every $\sigma$--fragment.
\end{theorem}
\begin{proof}
    The maximum of the Lorenz curve of a state $\rho$ is independent of the $\sigma$--fragment due to~\cref{lem:lcmax}, and can be expressed as an increasing function of mana,
    \begin{equation}
        \max_{x\in[0,1]}{\lc{\rho}{\sigma}(x)} = 1 + \sn{\rho} = \frac{1}{2} \left( 1 + e^\mana{\rho} \right).
    \end{equation}
    Therefore, the majorisation bound
    \begin{equation}
    	\lc{\rho}{\sigma}(x) \geq \lc{\tau}{\sigma}(x),\ x\in[0,1]
    \end{equation}
    implies the order $\max_{x\in[0,1]}{\lc{\rho}{\sigma}(x)} \geq \max_{x\in[0,1]}{\lc{\tau}{\sigma}(x)}$, hence the mana condition $\mana{\rho} \geq \mana{\tau}$.
\end{proof}

In sections~\cref{sec:unital,sec:stab}, we study majorisation constraints on magic distillation arising in different fragments. 
For this reason, we define the $n$-copy, $\epsilon$-noisy Strange state,
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)^{\otimes n} \coloneqq \left[ (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id \right]^{\otimes n},
\end{equation}
in $\O_\sigma$, where the pure magic state $\ket{\rm{S}}$ is induced with depolarising noise. 
Its Wigner distribution is visualised in~\cref{fig:strange}.

@@

In general, noisy circuits are well-described by the unital fragment.
To see this, consider incorporating noisy channels in the circuit, for example dephasing channels as in~\cref{eq:dephase} defined in different bases.
This process destroys the circuit symmetries, except for the invariance of the maximally mixed state.
Dephasing and bit-flip error channels are examples of the many error-inducing channels that respect the unital symmetry. \nick{Expand on significance of unital fragment}

We now consider the task of purifying $n$ copies of a noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ as given in~\cref{eq:noisysn} into a smaller number of copies $n'$ of a less noisy strange state $\rho_{\rm{S}}(\epsilon')^{\otimes n'}$, with $\epsilon' < \epsilon$ and $n' \leq n$,
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n-n')},
\end{equation}
where all copies $n, n', n - n'$ are even.
Since the state $\id / 3$ is free, tensoring in copies of it does not affect the distillation process.

The Lorenz curve $\lc{\rho_{\rm{S}}}{\id/3}$, for some general noise parameter $\epsilon$ and number of copies $n$, is defined at $9^n$ points between $0$ and $1$.
The exact expressions for the coordinates of these points can take $8$ different forms, depending on whether the noise level \nick{noise level / error rate} $\epsilon$ is greater or less than $\frac{3}{7}$, the parity of the number of copies $n$ is even or odd and the location relative to the curve peak is on the left hand side (LHS -- including the curve peak) or right hand side (RHS) of the curve peak.
The full details for the construction of all Lorenz curve forms are provided in~\cref{app:lcsu_technical}.

\nick{In the following, should we talk about the coordinates of all elbows and then say that we are gonna use only the first elbow, or should we leave full-elbow analysis for the appendix and only talk about first elbow here?}

Here we focus on the case of Lorenz curves with even copies $n, n'$ and low noise levels ($\epsilon' < \epsilon \leq 3/7$).
The part of the curve on the LHS of its peak is constructed by the positive components of the state Wigner distribution that take values $w_i$, with associated multiplicities $m_i$,
\begin{align}
	w_i &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2i}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{n-2i}, \\
	m_i &= 8^{2i}\binom{n}{2i},
\end{align}
where $i=0,\dots,\frac{n}{2}$.
It is readily seen that $\lc{\rho_{\rm{S}}}{\id/3}(x)$ has elbows on the LHS of its peak located at points $x=x_i$, where
\begin{equation}
	x_i = \Phi_+ \left( 2i; n , \frac{8}{9} \right),
\end{equation}
with $i = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
	\Phi_+(2i; n, p) \coloneqq \sum_{\ell=0}^{i} \binom{n}{2\ell} p ^{2\ell} (1-p)^{n-2\ell}.
\end{equation}
We provide technical details on the function $\Phi_+$ in the appendices.

The value of the Lorenz curve at an elbow is given by
\begin{equation}
	\lc{\rho}{\id/3}(x_{i}) \equiv L_{i} =  \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{n} \Phi_+\left(2i;n,\frac{12-4\epsilon}{15-8\epsilon}\right),
\end{equation}
and we define $(x_{-1}, L_{-1}) = (0,0)$.

\iffalse
The coordinates of any point on the Lorenz curve LHS can be expressed as convex sums of the neighbouring elbows,
\begin{align}
    x_{ijk} &= \left( 1-p_{ijk}\right) x_{i-1} + p_{ijk} x_{i} \label{eq:lcsu_xcoord}\\
    L_{ijk} &= \left( 1-p_{ijk} \right) L_{i-1} + p_{ijk} L_{i}, \label{eq:lcsu_lcoord}
\end{align}
where we use a three-index parametrisation for the coordinates, with $i$ the index of the elbow that follows, $j = 1,\dots,m_i$ and $k=1,\dots,9^{2(n-n')}$.
Therefore, for the initial state $\rho_{\rm{S}}^{\otimes 2n}$ we have
\begin{equation}
	p_{ijk} = \frac{j}{8^{2i}\binom{2n}{2i}},
\end{equation}
with $i=0,\dots,n,\ j(i) = 1,\dots,8^{2i}\binom{2n}{2i}$ and $k = 1$.
For the target state $\rho_{\rm{S}}(\epsilon')^{\otimes 2n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes 2(n-n')}$,
\begin{equation}
	p'_{i'j'k'} = \frac{k' + (j'-1)9^{2(n-n')}}{9^{2(n-n')} 8^{2i'}\binom{2n'}{2i'}}
\end{equation}
with $i'=0,\dots,n',\ j'(i') = 1,\dots,8^{2i'}\binom{n'}{2i'}$ and $k' = 1,\dots,9^{2(n-n')}$.
\fi

Lorenz curve comparison provides $9^n$ necessary constraints for the process~\cref{eq:sudist}.
\cref{thm:elbows} shows that they can be reduced to $n'$ independent constraints.
However, in principle we can pick any location $x_\star \in [0,1]$ at which the Lorenz curves are defined and we then get a necessary distillation constraint, 
\begin{equation}
	\lc{\rho_{\rm{S}}(n, \epsilon, 0)}{\frac{\id}{3}}(x_\star) \geq \lc{\rho_{\rm{S}}(n', \epsilon', n-n')}{\frac{\id}{3}}(x_\star).
\end{equation}

We choose $x_\star = x_0 = \Phi_+(0, n, \frac{8}{9}) = \frac{1}{9^{n}}$, the location of the first elbow of the initial state $\rho_{\rm{S}}(n, \epsilon, 0)$.
The initial state Lorenz curve coordinate at this location is
\begin{equation}
	L_\star = L_0 = \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{n} \Phi_+\left(0;n,\frac{12-4\epsilon}{15-8\epsilon}\right) = \left(\frac{3-4\epsilon}{9}\right)^n
\end{equation}

The first elbow of the target state is located at $x'_0 = \frac{1}{9^{n'}} > x_\star$, so in order to find the target state Lorenz curve coordinate $L'_\star$ at location $x_\star$, we can interpolate between the origin and the first target state elbow, 
\begin{equation}\label{eq:constraint_deriv}
	\begin{split}
	L'_\star &= \frac{x_\star}{x'_0}L'_0 = \frac{1}{9^{n-n'}}\left( \frac{5}{3} - \frac{8}{9}\epsilon'\ \right)^{n'}\Phi_+\left(0;n',\frac{12-4\epsilon'}{15-8\epsilon'}\right) \\
	&= \frac{(3-4\epsilon')^{n'}}{9^{n}}.
	\end{split}
\end{equation}
The process in~\cref{eq:sudist} is therefore possible only if $L_\star \geq L'_\star$ or equivalently,
\begin{align}
	\frac{n}{n'} &\geq \frac{\ln{(3-4\epsilon')}}{\ln{(3-4\epsilon)}}, \text{ or}\\
	\epsilon &\leq \frac{3}{4} - \frac{1}{4}(3-4\epsilon')^{\frac{n'}{n}}.
\end{align}
As an example, the process with $(n, n') = (12, 4)$ and $\epsilon' = 0.05$ yields $\epsilon \leq 0.398$.
Numerical comparison of the curves with varying $\epsilon$ suggests that we require $\epsilon \leq 0.373$ for the process to be possible.

Moreover, for the problem of distilling pure magic states where $\epsilon'=0$ we obtain an upper bound on the unital fragment distillation rate $R=n'/n$ given by
\begin{equation}
R \leq 1 + \frac{\ln (1 - \frac{4}{3} \epsilon)}{\ln 3},
\end{equation}
which holds for all finite $n$. 
In comparison, the bound from mana gives 
\begin{equation}
	R_{\rm{mana}} \leq 1 + \frac{\ln \left(1 - \frac{8}{15}\epsilon \right)}{\ln \frac{5}{3}},
\end{equation}
a threshold which is looser than the one provided by majorisation for all $\epsilon > 0$ as illustrated in~\cref{fig:distill_bounds}.
For example, for $\epsilon = 0.4$ we obtain $R\le 0.31 $ while the mana bound gives $R_{\rm{mana}} \le 0.53$.
\begin{figure}[t]
    \centering
    \includegraphics[scale=0.5]{figs/distill_bounds.pdf}
    \caption{\textbf{Distillation rate in unital fragment.} Distillation rates obtained by mana and majorisation are plotted for $\epsilon' = 0$, up to noise levels $\epsilon_{\rm{max}} = \frac{3}{7}$.
    Majorisation consistently provides stricter rates as proven in~\cref{thm:bounds}. \nick{add thauma?}
    }
    \label{fig:distill_bounds}
\end{figure}

@@

Given this context, we now provide the following result on bounding the distillation rate $R = R(\epsilon, \epsilon', \beta) \coloneqq \frac{n'}{n}$ of the process in~\cref{eq:stdist}.
The bounds depend on the free energy $F_\beta$ of state $\sigma$,
\begin{equation}
	F_\beta \coloneqq \tr[H \sigma] - \beta^{-1}S(\sigma) = -\beta^{-1}\log{\Z_\beta},
\end{equation}
where the von Neumann entropy is $S(\sigma) \coloneqq -\tr[\sigma\log{\sigma}]$.

\begin{theorem}\label{thm:stab_bounds}
	Let $\sigma$ be a qutrit stabiliser state, given by $\sigma = \frac{1}{\Z} e^{-\beta H}$ with $H$ having eigenvalues $E_0 \le E_1 \le E_2$, and where $\beta$ is an (effective) inverse temperature for the state. 
We define
\begin{equation}
	\beta_\star \coloneqq \frac{1}{E_2 - E_0} \ln{2}
\end{equation}
and for $\beta \leq \beta_\star$, we define a threshold error rate
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 3 - \dfrac{18}{8-e^{(E_2 - E_0)\beta}}.
\end{equation}
Then any distillation rate $R(\epsilon, \beta)$ in the $\sigma$--fragment of a magic theory is bounded as follows:\\
If $\beta \leq \beta_{\star}$ and $\epsilon  \leq \epsilon_{\star}$,
\begin{equation}
	R(\epsilon,\beta) \leq 1 + \frac{\ln{\left( 1 - \frac{4}{3}\epsilon \right)}}{\beta (E_0 - F_\beta)}.
\end{equation}
If $\beta \leq \beta_{\star}$ and $\epsilon_{\star} < \epsilon$, then
\begin{equation}
	R(\epsilon, \beta) \le 1 + \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)} - (E_2 - E_0)(\beta_{\star} - \beta)}{\beta (E_0 - F_\beta)}.
\end{equation}
Otherwise if $\beta > \beta_{\star}$,
\begin{equation}
	R(\epsilon, \beta) \leq  1+ \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)}}{-\ln{2} + \beta (E_2 - F_\beta)}.
\end{equation}
\end{theorem}
A few comments can be given on this result. 
Firstly, the specific numerical factors in $\epsilon_\star$ are a result of our choice of magic state. 
Secondly, these bounds are derived based on analysis of only a part of the Lorenz curves and can be improved via a finer analysis. 
This is apparent by simple numerical calculations on the entirety of the curves.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure. 
Finally, it is striking that we obtain a Landauer-like condition with a characteristic temperature $kT_\star \ln 2 = E_2 - E_0$, where $kT_\star = \beta_\star^{-1}$. 
It is unclear whether this points to a generic feature that can be directly related to fundamental thermodynamic relations, such as the erasure cost of a single bit being $kT \ln 2$. 
Given that we work with qutrits this seems surprising, but does deserve further study.

@@

\begin{proof}
	We provide an outline of the proof here, with all technical details included in~\cref{app:lcst_technical}.\\

The Wigner distribution of the $\epsilon$-noisy Strange state can be written as 
\begin{equation}
	\W[\bmx]{\rho_{\rm{S}}(\epsilon)} = (1-\epsilon)\W[\bmx]{\ketbra{\rm{S}}} + \epsilon\W[\bmx]{\frac{1}{3}\id},
\end{equation}
so we get positive components
\begin{equation}
	u(\epsilon) = \frac{1}{6} -\frac{1}{18}\epsilon
\end{equation}
at the 8 phase space points $\bmx \in \cal{P}_3 \setminus \{\bmo\}$ and a negative component
\begin{equation}
	- v(\epsilon) = - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right)
\end{equation}
at the origin $\bmx = \bmo$.

@@

Our aim is to obtain a distillation bound for~\cref{eq:stdist} which depends on variables $n, n', \epsilon, \epsilon'$ as well as $\beta$.
In the analysis that follows, we drop obvious variable dependencies for clarity.

>>>>

One of the most fundamental and commonly used magic monotones is the \emph{mana} of a state~\cite{cit:veitch2}, defined as
\begin{equation}
    \mana{\rho} \coloneqq \ln{(2\hspace{1pt}\sn{\rho}+1)},
\end{equation}
where the \emph{sum-negativity}~\cite{cit:veitch2} is the sum of the negative components in $\W{\rho}$,
\begin{equation}
    \sn{\rho} \coloneqq \sum\limits_{\bmx: \W[\bmx]{\rho} < 0} \abs{\W[\bmx]{\rho}}.
\end{equation}
Mana is an additive\footnote{It satisfies the condition $\mana{\rho_1 \otimes \rho_2} = \mana{\rho_1} + \mana{\rho_2}$ which is practical in distillation settings.} magic monotone, so it provides an analytical, necessary condition for many-copy magic state interconvertibility.

@@

Note that the Wigner components of an $n$-copy state $\rho^{\otimes n}$ can be calculated directly from $\W{\rho}$ by convolution of the distribution with itself,
\begin{equation}
	\W{\rho^{\otimes n}} = \W{\rho}^{\otimes n},
\end{equation}
where $\otimes$ can be interpreted as the usual Kronecker product in the last expression.
We may use the vector notation $\bmw(\rho) = (w(\rho)_i)_{i=1,\dots,d^2}$ for the Wigner distribution, in which case we can express the Kronecker product as
\begin{equation}
	\W{\rho} \otimes \W{\rho} = (w(\rho)_i w(\rho)_j)_{i,j=1,\dots,d^2}.
\end{equation}
The correspondence between the phase space and pure vector representations of the Wigner distribuition is discussed more in~\cref{app:cmpairs}.
\nick{In the rest of the paper we will be using the two notations interchangeably.}

@@

Normalisation of the Wigner distribution ensures that for all quantum states $\rho$, $\lc{\rho}{\sigma}(x) \geq 0$ and $\lc{\rho}{\sigma}(1) = 1$.
We stress that $0 \leq \lc{\rho}{\sigma}(x) \leq 1$ for all $x \in [0,1]$ if and only if $\rho$ is a positive Wigner state.
As a consequence, checking whether a magic state conversion of the form
\begin{equation}\label{eq:sigma_conv}
	\rho \xrightarrow{\E \in \O_\sigma} \tau
\end{equation} 
is not possible, reduces to the set of constraints
\begin{equation}\label{eq:majbound}
    \lc{\rho}{\sigma}(x) \geq \lc{\tau}{\sigma}(x),\ x\in [0,1],
\end{equation}
due to~\cref{thm:sigmamajor,thm:dmajor}.

We can refine the number of independent constraints stemming from this inequality. 
In fact, there are only as many independent constraints as there are elbows in the Lorenz curve of the target state as shown in~\cref{app:lc_constraints}.
However, in principle any one location $x$ provides a valid constraint leading to some upper distillation bound, while optimising over the location would provide the strictest bound.
In particular, we establish the \emph{first elbow constraint} which we use in~\cref{sec:unital,sec:stab} explicitly.
\begin{lemma}
	Consider a magic state interconversion as in~\cref{eq:sigma_conv}, where we denote by $(x_0, L_0)$ and $(x'_0, L'_0)$ the first elbow coordinates of the initial and target states respectively.
	If $x_0 < x'_0$, then it holds that
\begin{equation}\label{eq:first_elb_bound}
	\frac{L_0}{x_0} \geq \frac{L_0'}{x_0'}.
\end{equation}
\end{lemma}
\begin{proof}
	Consider the Lorenz curve constraint at $x = x_0$,
\begin{equation}
	\lc{\rho}{\sigma}(x_0) \geq \lc{\tau}{\sigma}(x_0).
\end{equation}
Since $x_0 < x'_0$, we can find the target state Lorenz curve coordinate $L'_\star$ at location $x = x_0$ by interpolating between the origin and the target state's first elbow, 
\begin{equation}\label{eq:constraint_deriv_temp}
	L'_\star = \frac{x_0}{x'_0}L'_0.
\end{equation}
We need $L_0 \geq L_\star'$ and rearranging we get~\cref{eq:constraint_deriv_temp}.
\end{proof}

We can now turn the attention from the first elbow to the peak of the Lorenz curve to associate it directly with the magic monotone of mana, and equivalently the sum-negativity of a quantum state. 
This holds independently of the particular $\sigma$--fragment one works in and as a result it becomes apparent that mana provides a weaker condition than majorisation for all magic state interconversions.

@@

We can therefore view mana as just one feature of the Lorenz curve, namely its maximum value. 
Conversely, it is now clear that the maximum of the Lorenz curve acts as a valid magic monotone.
\begin{theorem}\label{thm:bounds}
    Given a magic state conversion $\rho \longrightarrow \tau$, the majorisation condition is stronger than the mana condition in every $\sigma$--fragment.
\end{theorem}
\begin{proof}
    The maximum of the Lorenz curve of a state $\rho$ is independent of the $\sigma$--fragment due to~\cref{lem:lcmax}, and can be expressed as an increasing function of mana,
    \begin{equation}
        \max_{x\in[0,1]}{\lc{\rho}{\sigma}(x)} = 1 + \sn{\rho} = \frac{1}{2} \left( 1 + e^\mana{\rho} \right).
    \end{equation}
    Therefore, the majorisation bound
    \begin{equation}
    	\lc{\rho}{\sigma}(x) \geq \lc{\tau}{\sigma}(x),\ x\in[0,1]
    \end{equation}
    implies the order $\max_{x\in[0,1]}{\lc{\rho}{\sigma}(x)} \geq \max_{x\in[0,1]}{\lc{\tau}{\sigma}(x)}$, hence the mana condition $\mana{\rho} \geq \mana{\tau}$.
\end{proof}

In~\cref{sec:unital,sec:stab}, we study majorisation constraints on magic distillation arising in different fragments. 
For this reason, we define the $n$-copy, $\epsilon$-noisy Strange state,
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)^{\otimes n} \coloneqq \left[ (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id \right]^{\otimes n},
\end{equation}
in $\O_\sigma$, where the pure magic state $\ket{\rm{S}}$ is induced with depolarising noise. 
Its Wigner distribution is visualised in~\cref{fig:strange}.

We refer to parameter $\epsilon$ as \emph{noise level} to avoid confusion with the error rate $\delta$, a term commonly used in the literature to denote that the distilled state has a marginal overlap of at least $1-\delta$ with the desired magic state.
In our case the error rate of the state in~\cref{eq:noisysn} would be $\delta = \frac{2}{3}\epsilon$.

@@

In general, noisy circuits are well-described by the unital fragment.
To see this, consider incorporating noisy channels in the circuit, for example dephasing channels as in~\cref{eq:dephase} defined in different bases.
This process destroys the circuit symmetries, except for the invariance of the maximally mixed state.
Dephasing and bit-flip error channels are examples of the many error-inducing channels that respect the unital symmetry. \nick{Expand on significance of unital fragment}
\null\\

We consider the task of purifying $n$ copies of a noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ as given in~\cref{eq:noisysn} into a smaller number of copies $n'$ of a less noisy strange state $\rho_{\rm{S}}(\epsilon')^{\otimes n'}$, with $\epsilon' < \epsilon$ and $n' \leq n$,
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n-n')},
\end{equation}
where all copies $n, n', n - n'$ are even.
Since the state $\id / 3$ is free, tensoring in copies of it does not affect the distillation process.
The distillation rate $R \coloneqq n'/n$ for this process will in general depend on the noise levels, $R = R(\epsilon, \epsilon')$ and our task is to provide it with an upper bound.

The Lorenz curve $\lc{\rho_{\rm{S}}}{\id/3}$, for some general noise parameter $\epsilon$ and number of copies $n$, is defined at $9^n$ points between $0$ and $1$.
The exact expressions for the coordinates of these points can take $8$ different forms, depending on whether the noise level $\epsilon$ is greater or less than $\frac{3}{7}$, the parity of the number of copies $n$ is even or odd and the location relative to the curve peak is on the left hand side (LHS -- including the curve peak) or right hand side (RHS) of the curve peak.
The full details for the construction of all Lorenz curve forms are provided in~\cref{app:lcsu_technical}.

Here we focus on the comparison of the first elbow (which lies in the LHS part) of Lorenz curves with even copies $n, n'$ and low noise levels ($\epsilon' < \epsilon \leq 3/7$).

The Wigner distribution of the 1-copy, $\epsilon$-noisy Strange state can be written as 
\begin{equation}
	\W[\bmx]{\rho_{\rm{S}}(\epsilon)} = (1-\epsilon)\W[\bmx]{\ketbra{\rm{S}}} + \epsilon\W[\bmx]{\frac{1}{3}\id},
\end{equation}
so we get positive components
\begin{equation}
	u(\epsilon) \coloneqq \frac{1}{6} -\frac{1}{18}\epsilon
\end{equation}
at the 8 phase space points $\bmx \in \cal{P}_3 \setminus \{\bmo\}$ and a negative component
\begin{equation}
	- v(\epsilon) \coloneqq - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right)
\end{equation}
at the origin $\bmx = \bmo$.

For the sake of clarity, we may omit obvious variable dependencies in the following analysis.

The rescaled distribution in the unital fragment simply is 
\begin{equation}
	\widetilde{\rm{W}}_{\rho_{\rm{S}}|\frac{1}{3}\id}(\bmx) = \frac{\W[\bmx]{\rho_{\rm{S}}}}{\W[\bmx]{\frac{1}{3}\id}} = d\W[\bmx]{\rho_{\rm{S}}},
\end{equation}
so ordering the rescaled distribution is equivalent to ordering $\W{\rho_{\rm{S}}}$.

The component values and multiplicities ($m_i$) in the $n$-copy case are 
\begin{align}
	m_i &= 8^{2i}\binom{n}{2i}, \\
	w(\rho_{\rm{S}})_i &= u^{2i}(-v)^{n-2i}, \\
	w(\id/3)_i &=  \frac{1}{9^n},
\end{align}
where index $i$ runs through $0,\dots,\frac{n}{2}$.
This is derived in~\cref{app:lcsu_coord_elb}.

It is readily seen that the maximum Wigner component is achieved when $i=0$, so the corresponding multiplicity and Wigner component are $m_0 = 1$, $w(\rho_{\rm{S}})_0 = v^n$.
The first elbow coordinates therefore can be expressed as
\begin{equation}\label{eq:first_elb_coords}
	(x_0, L_0) = (m_0\hspace{2pt} w(\id/3)_0,\ m_0\hspace{2pt} w(\rho_{\rm{S}})_0) = \left(\frac{1}{9^n}, v^n \right)
\end{equation}

The first elbow of the target state is located at $x'_0 = \frac{1}{9^{n'}} > x_0$, so we can use the first elbow condition,
\begin{equation}\label{eq:first_elb_bound}
	\frac{L_0}{x_0} \geq \frac{L_0'}{x_0'},
\end{equation}
to compute analytical distillation bounds for distillation rate $R = R(\epsilon, \epsilon') \coloneqq n'/n$ in the unital fragment.
We substitute coordinates from~\cref{eq:first_elb_coords} appropriately in~\cref{eq:first_elb_bound} to get
\begin{equation}
	R \leq \frac{\ln{(3-4\epsilon)}}{\ln{(3-4\epsilon')}}.
\end{equation}

Specifically for the problem of distilling pure magic states ($\epsilon'=0$), we obtain an upper bound in the unital fragment given by
\begin{equation}
	R \leq 1 + \frac{\ln (1 - \frac{4}{3} \epsilon)}{\ln 3}.
\end{equation}

Similar upper bounds on distillation rates for qudits of odd prime dimension include the mana bound~\cite{cit:veitch} and the max--thauma bound~\cite{Wang2020} which is defined via a semi-definite program, but possesses properties that allow for an easy comparison with our bound.
The mana bound can be directly calculated as
\begin{equation}
	R \leq \frac{\mana{\rho_{\rm{S}}}(\epsilon)}{\mana{\rho_{\rm{S}}}(0)} = 1 + \frac{\ln \left(1 - \frac{8}{15}\epsilon \right)}{\ln \frac{5}{3}}.
\end{equation}
Using the properties of super-additivity and vanishing at free states of the max--thauma, we obtain
\begin{equation}
	\theta_{\rm{max}}(\rho_{\rm{S}}(\epsilon)) \geq (1-\epsilon) \ketbra{\rm{S}} + \epsilon \frac{1}{3}\id = (1-\epsilon) \theta_{\rm{max}}(\rho_{\rm{S}}(0)),
\end{equation}
so that for a deterministic single-copy distillation process, an initial number of copies $n \geq \theta_{\rm{max}}(\tau) / \theta_{\rm{max}}(\rho) \geq 1-\epsilon$ is required.
Therefore, the max--thauma bound for the pure Strange state distillation process can get as tight as
\begin{equation}
	R = \frac{1}{n} = \frac{\theta_{\rm{max}}(\rho_{\rm{S}}(\epsilon))}{\theta_{\rm{max}}(\rho_{\rm{S}}(0))} \leq 1-\epsilon.
\end{equation}

Both the mana and the max--thauma bounds are looser than the majorization bound we derived via the first elbow constraint as illustrated in~\cref{fig:distill_bounds}.
\begin{figure}[t]
    \centering
    \includegraphics[scale=0.5]{figs/distill_bounds.pdf}
    \caption{\textbf{Distillation bounds in the unital fragment.} Distillation bounds obtained by majorisation, mana and the line $1-\epsilon$ (tightest bound of max--thauma) are plotted for $\epsilon' = 0$, up to noise level $\epsilon = \frac{3}{7}$.
    Majorisation consistently provides stricter rates than mana and max--thauma.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

Given this context, we now provide the following result on bounding the distillation rate $R = R(\epsilon, \epsilon', \beta) \coloneqq \frac{n'}{n}$ of the process in~\cref{eq:stdist}.
We also write $R(\epsilon, \beta) = R(\epsilon, \epsilon'=0, \beta)$.
The bounds depend on the free energy $F_\beta$ of state $\sigma$,
\begin{equation}
	F_\beta \coloneqq \tr[H \sigma] - \beta^{-1}S(\sigma) = -\beta^{-1}\log{\Z_\beta},
\end{equation}
where the von Neumann entropy is $S(\sigma) \coloneqq -\tr[\sigma\log{\sigma}]$.

@@

\begin{theorem}\label{thm:stab_bounds}
	Let $\sigma$ be a qutrit stabiliser state, given by $\sigma = \frac{1}{\Z} e^{-\beta H}$ with $H$ having eigenvalues $E_0 \le E_1 \le E_2$, and where $\beta$ is an (effective) inverse temperature for the state. 
We define
\begin{equation}
	\beta_\star \coloneqq \frac{1}{E_2 - E_0} \ln{2}
\end{equation}
and for $\beta \leq \beta_\star$, we define a threshold noise level
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 3 - \dfrac{18}{8-e^{(E_2 - E_0)\beta}}.
\end{equation}
Then any distillation rate $R(\epsilon, \beta)$ in the $\sigma$--fragment of a magic theory is bounded as follows:\\
If $\beta \leq \beta_{\star}$ and $\epsilon  \leq \epsilon_{\star}$,
\begin{equation}
	R(\epsilon,\beta) \leq 1 + \frac{\ln{\left( 1 - \frac{4}{3}\epsilon \right)}}{\beta (E_0 - F_\beta)}.
\end{equation}
If $\beta \leq \beta_{\star}$ and $\epsilon_{\star} < \epsilon$, then
\begin{equation}
	R(\epsilon, \beta) \le 1 + \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)} - (E_2 - E_0)(\beta_{\star} - \beta)}{\beta (E_0 - F_\beta)}.
\end{equation}
Otherwise if $\beta > \beta_{\star}$,
\begin{equation}
	R(\epsilon, \beta) \leq  1+ \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)}}{-\ln{2} + \beta (E_2 - F_\beta)}.
\end{equation}
\end{theorem}
A few comments can be given on this result. 
Firstly, the specific numerical factors in $\epsilon_\star$ are a result of our choice of magic state. 
Secondly, these bounds are derived based on analysis of only a part of the Lorenz curves and can be improved via a finer analysis. 
This is apparent by simple numerical calculations on the entirety of the curves.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure. 
Finally, it is striking that we obtain a Landauer-like condition with a characteristic temperature $kT_\star \ln 2 = E_2 - E_0$, where $kT_\star = \beta_\star^{-1}$. 
It is unclear whether this points to a generic feature that can be directly related to fundamental thermodynamic relations, such as the erasure cost of a single bit being $kT \ln 2$. 
Given that we work with qutrits this seems surprising, but does deserve further study.

\begin{proof}
	We provide the full proof here, with some technical details pushed to~\cref{app:lcst_technical}.\\

@@

Our aim is to obtain a distillation bound for~\cref{eq:stdist} which depends on variables $n, n', \epsilon, \epsilon'$ as well as $\beta$.
In the analysis that follows, we again drop obvious variable dependencies for clarity.
